### 2013-09-18 draft.tex
Added abstract

\documentclass[english,12pt]{article}
\usepackage[T1]{fontenc}
\usepackage[latin9]{inputenc}
\usepackage{float}
\usepackage{graphicx}
\usepackage{esint}
\usepackage{natbib}
\usepackage{times,bm}
\usepackage{babel}

\makeatletter
\linespread{1.25}
\pagestyle{headings}

\begin{document}

\title{Implementation of Parallel Algorithms for ARIMA in Distributed
  Database Systems}

% \author{}

\maketitle

\begin{abstract}
  MADlib is an open-source library for scalable in-database
  analytics. It provides data-parallel implementations of
  mathematical, statistical and machine-learning methods for
  structured and unstructured data. We impemented ARIMA in MADlib's
  framework. The algorithms for fitting ARIMA model to time series are
  intrinsically sequential, because any calculation for a specific
  time $t$ depends on the
  result from the calculation for $t-1$. This makes it difficult to
  parallelize the ARIMA model fitting. Our solution parallelizes the
  computation by splitting the data into $n$ chunks. Since the model
  fitting involves multiple iterations of computations. We use the
  results from previous iteration as the initial values for each
  chunk. Thus the computation for each chunk of data does not need to
  wait for the results from previous chunk. Another technique is used
  to improve the performance of the implementation, which
  redistributes the original data so that each chunk can be completely
  loaded into memory.
\end{abstract}

\section{Introduction}

% Introduce to MADlib, and implement ARIMA in MADLib

\section{MADlib}

% More details on MADlib

\section{Implementation of ARIMA}

% A short summary for the next few subsections

\subsection{The Algorithm}

% List the LM algorithm

\subsection{The Problems in Parallelization}

% What are the difficulties that we are facing

\subsection{Our Solution}

% Our solution and how to improve the performance

\subsection{Performance}

% Good performance improvement, some examples

\subsection{Discussion}

% Can be applied to other algorithms

\section{Conclusion}

% \bibliographystyle{apalike}
% \bibliography{financial_market}

\end{document}