### 2013-09-19 draft.tex
Minor changes

\documentclass[english, 12pt]{article} \usepackage[T1]{fontenc}
\usepackage[latin9]{inputenc} \usepackage{float} \usepackage{graphicx}
\usepackage{esint} \usepackage{natbib} \usepackage{times,bm} \usepackage{babel}

\makeatletter
\linespread{1.25}
\pagestyle{headings}

\begin{document}

\title{Implementation of Parallel Algorithms for ARIMA in Distributed
  Database Systems}

% \author{}

\maketitle

\begin{abstract} 
    MADlib is an open-source library for scalable in-database analytics. We
    impemented ARIMA in MADlib's framework. The algorithms for fitting ARIMA
    model to time series are intrinsically sequential, because any calculation
    for a specific time $t$ depends on the result from the calculation for
    $t-1$.  This makes it difficult to parallelize the ARIMA model fitting. Our
    solution parallelizes the computation by splitting the data into $n$
    chunks. Since the model fitting involves multiple iterations of
    computations. We use the results from previous iteration as the initial
    values for each chunk. Thus the computation for each chunk of data does not
    need to wait for the results from previous chunk. Another technique is used
    to improve the performance of the implementation, which redistributes the
    original data so that each chunk can be completely loaded into memory.
\end{abstract}

\section{Introduction}

% Introduce to MADlib, and implement ARIMA in MADLib
MADlib \cite{madlib} is an open source library for scalable in-database
analytics.  It is created by the Predictive Analytics Team at Pivotal Inc.
MADlib provides data-parallel implementations of mathematical, statistical and
machine-learning methods for structured and unstructured data.

\section{MADlib}

% More details on MADlib

\section{Implementation of ARIMA}

% A short summary for the next few subsections

\subsection{The Algorithm}

% List the LM algorithm

\subsection{The Problems in Parallelization}

% What are the difficulties that we are facing

\subsection{Our Solution}

% Our solution and how to improve the performance

\subsection{Performance}

% Good performance improvement, some examples

\subsection{Discussion}

% Can be applied to other algorithms

\section{Conclusion}

\bibliographystyle{apalike}

\bibliography{pred}

\end{document}